Financialization of Commodities: an Asset Pricing Perspective

### 2024-06-11 others/paper/elsevier article/main.Rmd
Track 'others' folder

---
documentclass: article
# classoption: letterpaper
geometry: "letterpaper, left = 1in, right = 1in, top = 1.5in, bottom = 1.5in"
link-citations: TRUE
bibliography: BibTeX.bib
biblio-style: jf
fontsize: 12pt

output:
  pdf_document:
    template: "template.tex"
    citation: natbib
    fig_caption: TRUE
    keep_tex: TRUE
    number_sections: TRUE
    latex_engine: pdflatex
    includes:
     in_header: preamble.tex

title: 'Financialization of Commodities: an Asset Pricing Perspective'

author: 
  - name: Devraj Basu
    surname: Basu
    email: devraj.basu@strath.ac.uk
    affiliation: University of Strathclyde, Glasgow, UK
  - name: Olivier Bauthéac
    surname: Bauthéac
    email: olivier.bautheac@strath.ac.uk
    affiliation: University of Strathclyde, Glasgow, UK

jel: G11, G23, G13, Q02
keywords: Commodity futures, financialization, hedging pressure, financial crisis

@@

---

```{r, echo = FALSE, message = FALSE}
packages <- c("tidyr", "magrittr", "knitr", "lettercase", "tinytex", "bookdown", "kableExtra", "here")
invisible(sapply(packages, require, character.only = TRUE))
options(knitr.table.format = "latex")
knitr::opts_chunk$set(echo = FALSE, results = "asis")
tickers <- readRDS(file = file.path(here::here(), "data", "tickers.rds"))
tables <- readRDS(file = file.path(here::here(), "explore", "results", "tables.rds.rds"))

@@

* Devraj Basu: 
  I am a co-author of this paper and wish to state that I have no conflict of interest, financial or otherwise. I do not hold paid or unpaid positions at any organization whose goals relate to the article.  
  ![](pics/signature - Devraj.png)\

@@

* Olivier Bauthéac:
  I am a co-author of this paper and wish to state that I have no conflict of interest, financial or otherwise. I do not hold paid or unpaid positions at any organization whose goals relate to the article.  
  ![](pics/signature - Olivier.png)\

@@

# Introduction

@@

The cross-sectional nature of financialization [@basak_model_2016; @cheng_financialization_2014] leads us to extend the Keynesian individual hedging pressure paradigm by incorporating aggregate CHP (\textbf{CHP}), a market wide measure of hedging pressure first considered in @hong_what_2012. This paradigm implies that returns for individual commodities should be high in periods where \textbf{CHP} is low (\textbf{backwardation}) and low in periods where \textbf{CHP} is high (\textbf{contango}). Over the first period the cross sectional average returns during \textbf{backwardation} for both the US and UK commodities are positive with those for the US statistically significant, and significantly higher than during \textbf{contango}, providing broad support for an aggregate version of the Keynesian hedging pressure hypothesis. Over the second period this pattern reverses, with both the US and UK commodities having significant positive returns during \textbf{contango} with these returns being significantly higher than during \textbf{backwardation}. This change in pattern suggests that the onset of financialization altered the traditional risk shifting nature of hedger’s behaviour, which is implied by the Keynesian hypothesis. The arrival of financial investors raises the question of whether they functioned as liquidity providers or liquidity demanders. Arguments have been made for both of these possibilities with @moskowitz_time_2012 arguing that financial investors provide liquidity while @Kangtaletwopremiums2017 argue that they demand liquidity. Our analysis of returns in periods of low and high aggregate CHP before and during financialization allows us to analyse this issue from a new perspective and suggests that financial investors may have been liquidity demanders leading hedgers to become liquidity providers.

@@

The onset of the financial crisis and the monetary policy regimes that followed, on the other hand, also appear to have induced significant changes in pricing dynamics. Returns on both US and UK traded commodities fell dramatically over both the third and fourth periods with the pattern of return for the US traded commodities across \textbf{CHP} regimes reverting to Keynesian paradigm. The pricing results indicate the emergence of a systematic factor across the entire cross-section of the US commodities as well as strong evidence of cross-market linkages.

Financialization was an issue of such policy importance that it triggered legislative action. The debate was initially framed around adequacy of speculation, the burning issue of the day, and the consequent analysis focused on the more mechanical effects of financialization. With the benefit of hindsight, it appears this approach was perhaps too narrow and it now seems necessary to address financialization from a broader perspective. Commodity price dynamics appear to have altered substantially in quite different ways over the financialization and the financial crisis periods and our commodity futures based asset pricing approach seems able to provide new insights into the nature of these changes: financialization was a phenomenon endogenous to the commodity markets transmitted via the commodity futures markets and mainly concentrated in a specific sector (metals) while the crisis and its aftermath seems to have delivered an exogenous shock across the entire cross-section of global commodities. The initial view of financialization was that it consisted of speculative flows which had the effect of driving up and creating bubble like conditions in commodity future prices^[@de_schutter_food_2010_1; @gilbert_how_2010; @gilbert_speculative_2010; @masters_testimony_2008; @masters_accidental_2008; @herman_not_2011_1; @schumann_hunger-makers_2011; @singleton_investor_2013; @unctad_global_2009.]. However, the fundamental question of the nature of the impact of financialization across the entire cross-section of commodity futures markets has not yet been completely answered. We provide an empirical complement to a new stream of theoretical studies that try to model the impact of financialization on various aspects of commodity futures markets^[@etula_broker-dealer_2013, @acharya_limits_2013, @cheng_convective_2014, @leclercq_equilibrium_2014, @sockin_informational_2015, @goldstein_speculation_2013, @EkelandSpeculationcommodityfutures2016, @goldstein_commodity_2017.] and demonstrate that the effects of financialization extend beyond the mechanical effects induced by indexation as outlined in @basak_model_2016 as well as @tang_index_2012.

@@

A world-wide debate has ensued about the role of index funds in commodity markets. The first responses to the 2007/2008 crisis of escalating food and energy prices took the form of policy reports, many of which reasoned that the growth of commodity index funds came along with an influx of largely speculative capital that was responsible for driving commodity prices beyond their historic highs [@de_schutter_food_2010_1; @gilbert_speculative_2010; @herman_not_2011_1; @schumann_hunger-makers_2011; @unctad_global_2009]. The US and various european countries also became involved in this issue. Early US Senate investigations drew on pricing and trading data supplied by the CFTC as well as interviews with numerous experts including hedge fund manager Michael Masters who linked the growing presence of index investors in US wheat markets to the observed surges in both futures and cash prices [@masters_testimony_2008; @masters_accidental_2008]). This contention has since come to be known as the Masters’ hypothesis and was largely endorsed by the final US Senate report on the issue [@USSenate_Excessive_speculation_wheat_2009]. At this stage, the question thrust on the academic sphere was whether “excessive speculation” (understood as the market activity peculiar to index type investors) was linked to escalating energy and agricultural commodity prices. A number of academic studies, most of which used commodity specific causality and correlation based analysis have disproved this contention, particularly in the case of agricultural and energy commodities. @irwin_devil_2009, @sanders_adequacy_2010, @sanders_impact_2011, @sanders_new_2011, @irwin_commodity_2013, @brunetti_commodity_2014, @hamilton_effects_2015 and @bruno_financialization_2017 study the agricultural markets. @buyuksahin_speculators_2011, @tokic_speculation_2012, @fattouh_role_2013, @kilian_role_2014, @knittel_simple_2016 and @manera_modelling_2016 study the energy markets. @bohl_does_2013_2, @kim_does_2015 and @boyd_prevalence_2016 study both energy and agricultural markets while @irwin_index_2011, @irwin_financialization_2012 and @stoll_commodity_2011 study the commodity markets in general.

@@

Commodity index funds (CIFs), commodity based exchange traded funds (ETFs) and notes (ETNs) as well as commodity linked notes (CLNs) were amongst the main investment vehicules that allowed institutional and retail investors to build up commodity exposure in the early 2000s [@boons_stock_2012; @henderson_new_2015; @irwin_index_2011; @unctad_global_2009; @schumann_hunger-makers_2011]. Most of these products were primarily designed to track prominent commodity indexes amongst which the S&P-GSCI, the most popular. Most of its constituents at the time of writting form the basis of the broad cross-section of commodities that we consider in this study^[See section \ref{assets} for details on the commodity pool considered.]. Unleaded gasoline, the NYMEX legacy contract for gasoline stopped trading in 2006 when it was replaced by Reformulated Gasoline Blendstock for Oxygen Blending (RBOB) with the two futures series trading alongside for most of the year. For our NYMEX gasoline data series, we consider unleaded gasoline up to September 1^st^, 2006 and RBOB gasoline thereafter, date at which liquidity for the latter overtook that for the former.

@@

On the one hand, futures markets have a long history of serving commodity producers to allievate their commodity price and output risks. @keynes_treatise_1930 and @hicks_value_1939 emphasise the "normal" behaviour of naturally short hedgers outsourcing their commodity risk to naturally long commodity specialist speculators. This market configuration is referred to as "backwardation" where futures is expected to trade at a discount to spot, providing speculators with incentives to take the long side. The opposite market paradigm is referred to as "contango". Working [@working_theory_1948; @working_hedging_1953] and Cootner [@cootner_returns_1960; @CootnerSpeculationhedging1967] somehow relax this "supply-of-speculative-services" [@Tilllongtermperspectivecommodity2007] approach by introducing processors and merchants, also commodity experts, on the hedgers side and relating the notions of backwardation and contango more directly to the term structure via the theory of storage where the shape of the front end of the term structure (basis) is directly related to the market price of storage.  
Financialization, on the other hand, refers to the entry of financial investors into the commodity futures markets who, for the most part, are not commodity experts. Besides, these new market participants tend to exhibit herding like behaviour, often taking massive long-only positions on the whole cross-section of indexed commodities in an attempt to hedge against inflation and/or seek further diversification benefits for portfolios with, in most cases, existing large positions across various asset classes [@brunetti_speculators_2016; @boyd_prevalence_2016; @cheng_convective_2014; @juvenal_speculation_2015; @singleton_investor_2013; @tang_index_2012]. This contrasts with the traditional approaches of Keynes and Working and in this context the issues of co-movement and aggregate market participants behaviour seem particularly relevant.  
Asset pricing techniques are well suited to study co-movement. Building on theory we develop an ad-hoc commodity based asset pricing framework that we believe is well adapted to study these two issues simultaneously and implement it over four periods of interest. The first, the pre-financialization period, starts in 1997 and is naturally followed by financialization, spanning the 2004/2007 period, and the financial crisis and its aftermath, spanning the 2008/2013 period. The last period spans the 2014/present period sarting with the beginning of the tapering/end of the US Quantitative Easing program (QE).  
Using the weekly price, open interest and market position data described above we construct market portfolios of commodity nearby futures as well as mimicking portfolios for market, commercial hedging pressure, term structure and liquidity risk factors in returns. While our market portfolios are long only, equally weighted portfolios of particular commodity sets, our other mimicking portfolios for risk factors are constructed as combinations of two sub-portfolios (legs), one held long, one held short, which respective constituents are pooled on a weekly basis according to risk factor specific criterias. The weekly return for a particular mimicking portfolio is calculated as the difference between the average return for constituents of the long portfolio and that for the constituents of the short portfolio. i.e, return on the mimicking portfolio for risk factor $i$ ($f_{i}$) for week $t$:

@@

The same operation is repeated for the two ($p=2$) and three ($p=3$) factor model cases, where regressors include in turn all combinations of any two and three risk factors respectivly, as well as for the four ($p=4$) factor model case where all the risk factor mimicking portfolios are included. Results are reported for the one-factor model case with the other results available from the authors upon request.  
For each period we also implement the analysis described above over periods of low and high aggregate CHP (\textbf{CHP}), a notion which we introduce in this study and which we believe is particularly relevant in the context of financialization where issues of co-movement and aggregate market participants' behaviour play a central role. We define \textbf{CHP} as the cross-sectional average of CHPs for a particular pool of assets. i.e, Week $t$ \textbf{CHP} for asset pool $x$:

@@

We further define aggregate backwardation (\textbf{backwardation}) and contango (\textbf{contango}) \textbf{CHP} regimes as, for a given period, \textbf{CHP} levels below and above the period's median respectively and, in contrast with [@hong_what_2012] who aggregate across sub-sectors, the asset pool we consider for \textbf{CHP} construction include the whole cross-section of US traded commodities.

@@

For each period and \textbf{CHP} regime we record the proportion of time each pick is held in the long and (/or) short leg of the corresponding risk factor mimicking portfolio as well as the average pairwise correlation amongst every sets of picks.

@@

# Results & discussion

The descriptive statistics over the various time periods for our two sets of commodities (US and non-US) as well as the four factors considered are shown in Table \ref{table1}. The traditional Keynesian hedging pressure paradigm postulates a negative relationship between the return on an individual commodity and its hedging pressure. Financialization has been seen as a cross-sectional commodity market phenomenon; this naturally leads us to extend the individual hedging pressure paradigm by incorporating aggregate CHP (\textbf{CHP}), a market wide measure of hedging pressure. The extended Keynesian hedging pressure paradigm should also imply that returns for individual commodities should be high in periods where \textbf{CHP} is low and vice et versa.  
In table \ref{table2} the mean returns are shown over high (\textbf{contango}) and low (\textbf{backwardation}) \textbf{CHP} regimes for each time period. Over the first period eighteen of the twenty-four US commodities have higher mean returns during periods of \textbf{backwardation} as compared to \textbf{contango}, with the difference being statistically significant at the 5% level for one commodity and for the remaining six commodities the higher mean return during \textbf{contango} is not statistically significant for any of them. For the six UK commodities, all metals, four of them had higher mean returns during phases of \textbf{backwardation} as compared to \textbf{contango} in the first period, with one of these differences being statistically significant at the 5% level while for the remaining two the higher mean return during \textbf{contango} is not statistically significant for either. An equally weighted portfolio of the twenty-four US commodities had a mean return of 10.98% during \textbf{backwardation} phases, statistically significant at the 1% level, and a mean return of 0.93% during \textbf{contango}, both in the first period, with the difference between them statistically significant at the 5% level. The corresponding figures for an equally weighted portfolio of the UK metals were 8.35% and -1.71%, the difference again statistically significant at the 5% level. These results provide broad support for an aggregate version of the Keynesian hedging pressure hypothesis and suggests that hedgers in the aggregate were engaged in risk transfer during this period.  
Over the second period the pattern essentially reverses. For the twenty-four US commodities over this period, eighteen have higher mean returns during phases of \textbf{backwardation} over \textbf{contango}, with the difference statistically significant for four commodities and of the remaining seven the higher mean return during \textbf{backwardation} is statistically significant for one commodity. All of the six UK metals achieved higher mean returns over \textbf{backwardation} than contango with the difference being statistically significant for one commodity. The equally weighted US portfolio had a mean return of 11.16% over \textbf{backwardation} while it was 26.68% during \textbf{contango}, the first statistically significant at the 5% level and the second at the 1% level, with the difference statistically significant at the 1% level. The difference was considerably more pronounced for the equally weighted UK metals portfolio with a mean return of 6.43% during \textbf{backwardation} rising to 41.22% during \textbf{contango}, with the difference also significant at the 1% level.  
Thus the onset of financialization seems to have engendered a change in the nature of hedger’s behaviour in the aggregate, taking it away from risk shifting, the traditional Keynesian view. This phenomenon has been noted in earlier theoretical and empirical work [@DanthineInformationfuturesprices1978, @stout_why_1998] in different context from which two competing models of behaviour have emerged. The information arbitrage model implies that hedgers may often be seeking out counterparties to trade with rather than being purely passive. This issue is also raised in both @cheng_financialization_2014 and @StulzRethinkingriskmanagement1996 who point out that hedgers may be taking a view on prices just as speculators do. In fact, by hedging away some of their risk, hedgers are able to speculate more heavily based on their disagreement against speculators regarding futures price movement [@SimsekSpeculationrisksharing2013], which fits into a heterogenous expectation theory of speculation. As @stout_why_1998 points out, this disagreement-based theory of trading is one of the main reason the public and the law disapproves the speculators as this form of trading is regarded as non-productive^[@DuffieChallengespolicytreatment2014 also discusses some of the challenges faced by a policy treatment of speculative trading motivated by differences in beliefs.]. In the second model hedgers are seen as liquidity providers as in @Kangtaletwopremiums2017. This explanation is also consistent with the nature of financialization which led to the arrival of long term, long only investors who require substantial liquidity to roll-over their positions.

The average time series pricing performance of the various one factor models on the US traded assets over the four periods is shown in Table \ref{table3}. Over the first period the market factor has the best average time series performance with the highest average adjusted R^2^ with all three futures market factors performing poorly and the overall pricing performance consistent across periods of high and low \textbf{CHP} (below period median). The two factor models average R^2^s (results available upon request) are roughly the sum of the one factor R^2^s indicating that the factors are almost orthogonal over this period. The onset of financialization leads to a substantial improvement in the performance of the CHP factor with its average R^2^ increasing to 5.4% in the second period relative to 1.4% in the first. The performance of the market factor also improves though not by as much in a relative sense (18.6% from 12.2%). Both the term structure factor and the open interest growth factor show a much smaller improvement in pricing performance. Over all periods, all of our factor models show similar pricing performance across regimes of high and low \textbf{CHP}.

@@

kable(table, align = c(rep('l', 2L), rep('r', 4L)), digits = getOption("digits"),
      booktabs = TRUE, longtable = TRUE, linesep = "", caption = caption) %>%
  kableExtra::kable_styling(full_width = TRUE, font_size = 10L) %>%
  kableExtra::column_spec(1L, width = "5cm") %>%
  kableExtra::column_spec(2L, width = "2cm")

@@

kable(table, align = c(rep('l', 3L), rep('r', 4L)), digits = getOption("digits"), 
      booktabs = TRUE, longtable = TRUE, linesep = "", caption = caption) %>% 
  kableExtra::kable_styling(font_size = 10L, full_width = TRUE) %>% 
  kableExtra::column_spec(1L, width = "3cm") %>% 
  kableExtra::column_spec(2L, width = "3cm") %>% 
  kableExtra::column_spec(3L, width = "2cm")
```

@@

kable(table, align = c(rep('l', 2L), rep('r', 4L)), digits = getOption("digits"), 
      booktabs = TRUE, longtable = TRUE, linesep = "", caption = caption) %>% 
  kableExtra::kable_styling(font_size = 10L, full_width = TRUE) %>% 
  kableExtra::column_spec(1L, width = "5cm") %>% 
  kableExtra::column_spec(2L, width = "2cm")
```

>>>>

---
title: "Financialization of Commodities: an Asset Pricing Perspective"
author:
  - name: Devraj Basu
    email: devraj.basu@strath.ac.uk
    affiliation: strathclyde
  - name: Olivier Bauthéac
    email: olivier.bautheac@strath.ac.uk
    affiliation: strathclyde
address:
  - code: strathclyde
    address: University of Strathclyde, Glasgow, UK

date: "`r Sys.Date()`"
csl: elsevier-harvard.csl
output: rticles::elsevier_article
# output: bookdown::pdf_document2
bibliography: "C:/Users/yyb14214/Dropbox/research/literature/references.bib"

@@

jel: G11, G23, G13, Q02
keywords: Commodity futures, financialization, hedging pressure, asset pricing
---

@@

* Devraj Basu: 
  I am a co-author of this paper and wish to state that I have no conflict of interest, financial or otherwise. I do not hold paid or unpaid positions at any organization whose goals relate to the article.  
  ![](Devraj.jpg)

@@

* Olivier Bauthéac:
  I am a co-author of this paper and wish to state that I have no conflict of interest, financial or otherwise. I do not hold paid or unpaid positions at any organization whose goals relate to the article.  
  ![](Olivier.jpg)

@@

```{r, echo = FALSE, message = FALSE, warning = FALSE}
packages <- c("tidyr", "magrittr", "knitr", "tinytex", "bookdown", "kableExtra", "here")
invisible(sapply(packages, require, character.only = TRUE))
options(knitr.table.format = "latex")
knitr::opts_chunk$set(echo = FALSE, results = "asis")
tickers <- readRDS(file = file.path(here::here(), "data", "tickers.rds"))
tables <- readRDS(file = file.path(here::here(), "explore", "results", "tables.rds"))

# Introduction {#intro}

@@

The cross-sectional nature of financialization [@basak_model_2016; @cheng_financialization_2014] leads us to extend the Keynesian individual hedging pressure paradigm by incorporating aggregate CHP (**CHP**), a market wide measure of hedging pressure first considered in @hong_what_2012. This paradigm implies that returns for individual commodities should be high in periods where **CHP** is low (**backwardation**) and low in periods where **CHP** is high (**contango**). Over the first period the cross sectional average returns during **backwardation** for both the US and UK commodities are positive with those for the US statistically significant, and significantly higher than during **contango**, providing broad support for an aggregate version of the Keynesian hedging pressure hypothesis. Over the second period this pattern reverses, with both the US and UK commodities having significant positive returns during **contango** with these returns being significantly higher than during **backwardation}. This change in pattern suggests that the onset of financialization altered the traditional risk shifting nature of hedger’s behaviour, which is implied by the Keynesian hypothesis. The arrival of financial investors raises the question of whether they functioned as liquidity providers or liquidity demanders. Arguments have been made for both of these possibilities with @moskowitz_time_2012 arguing that financial investors provide liquidity while @kang_tale_2017 argue that they demand liquidity. Our analysis of returns in periods of low and high aggregate CHP before and during financialization allows us to analyse this issue from a new perspective and suggests that financial investors may have been liquidity demanders leading hedgers to become liquidity providers.

@@

The onset of the financial crisis and the monetary policy regimes that followed, on the other hand, also appear to have induced significant changes in pricing dynamics. Returns on both US and UK traded commodities fell dramatically over both the third and fourth periods with the pattern of return for the US traded commodities across **CHP** regimes reverting to Keynesian paradigm. The pricing results indicate the emergence of a systematic factor across the entire cross-section of the US commodities as well as strong evidence of cross-market linkages.

Financialization was an issue of such policy importance that it triggered legislative action. The debate was initially framed around adequacy of speculation, the burning issue of the day, and the consequent analysis focused on the more mechanical effects of financialization. With the benefit of hindsight, it appears this approach was perhaps too narrow and it now seems necessary to address financialization from a broader perspective. Commodity price dynamics appear to have altered substantially in quite different ways over the financialization and the financial crisis periods and our commodity futures based asset pricing approach seems able to provide new insights into the nature of these changes: financialization was a phenomenon endogenous to the commodity markets transmitted via the commodity futures markets and mainly concentrated in a specific sector (metals) while the crisis and its aftermath seems to have delivered an exogenous shock across the entire cross-section of global commodities. The initial view of financialization was that it consisted of speculative flows which had the effect of driving up and creating bubble like conditions in commodity future prices^[@deschutter_food_2010; @gilbert_how_2010; @gilbert_speculative_2010; @masters_testimony_2008; @masters_accidental_2008; @herman_not_2011; @schumann_hunger_2011; @singleton_investor_2013; @unctad_global_2009.]. However, the fundamental question of the nature of the impact of financialization across the entire cross-section of commodity futures markets has not yet been completely answered. We provide an empirical complement to a new stream of theoretical studies that try to model the impact of financialization on various aspects of commodity futures markets^[@etula_broker-dealer_2013, @acharya_limits_2013, @cheng_convective_2014, @leclercq_equilibrium_2014, @sockin_informational_2015, @goldstein_speculation_2014, @ekeland_speculation_2016, @goldstein_commodity_2017.] and demonstrate that the effects of financialization extend beyond the mechanical effects induced by indexation as outlined in @basak_model_2016 as well as @tang_index_2012.

@@

A world-wide debate has ensued about the role of index funds in commodity markets. The first responses to the 2007/2008 crisis of escalating food and energy prices took the form of policy reports, many of which reasoned that the growth of commodity index funds came along with an influx of largely speculative capital that was responsible for driving commodity prices beyond their historic highs [@deschutter_food_2010; @gilbert_speculative_2010; @herman_not_2011; @schumann_hunger_2011; @unctad_global_2009]. The US and various european countries also became involved in this issue. Early US Senate investigations drew on pricing and trading data supplied by the CFTC as well as interviews with numerous experts including hedge fund manager Michael Masters who linked the growing presence of index investors in US wheat markets to the observed surges in both futures and cash prices [@masters_testimony_2008; @masters_accidental_2008]). This contention has since come to be known as the Masters’ hypothesis and was largely endorsed by the final US Senate report on the issue [@senate_excessive_2009]. At this stage, the question thrust on the academic sphere was whether “excessive speculation” (understood as the market activity peculiar to index type investors) was linked to escalating energy and agricultural commodity prices. A number of academic studies, most of which used commodity specific causality and correlation based analysis have disproved this contention, particularly in the case of agricultural and energy commodities. @irwin_devil_2009, @sanders_adequacy_2010, @sanders_impact_2011, @sanders_new_2011, @irwin_commodity_2013, @brunetti_commodity_2014, @hamilton_effects_2015 and @bruno_financialization_2017 study the agricultural markets. @buyuksahin_speculators_2011, @tokic_speculation_2012, @fattouh_role_2013, @kilian_role_2014, @knittel_simple_2016 and @manera_modelling_2016 study the energy markets. @bohl_does_2013, @kim_does_2015 and @boyd_prevalence_2016 study both energy and agricultural markets while @irwin_index_2011, @irwin_financialization_2012 and @stoll_commodity_2011 study the commodity markets in general.

@@

Commodity index funds (CIFs), commodity based exchange traded funds (ETFs) and notes (ETNs) as well as commodity linked notes (CLNs) were amongst the main investment vehicules that allowed institutional and retail investors to build up commodity exposure in the early 2000s [@boons_stock_2012; @henderson_new_2015; @irwin_index_2011; @unctad_global_2009; @schumann_hunger_2011]. Most of these products were primarily designed to track prominent commodity indexes amongst which the S&P-GSCI, the most popular. Most of its constituents at the time of writting form the basis of the broad cross-section of commodities that we consider in this study^[See section \ref{assets} for details on the commodity pool considered.]. Unleaded gasoline, the NYMEX legacy contract for gasoline stopped trading in 2006 when it was replaced by Reformulated Gasoline Blendstock for Oxygen Blending (RBOB) with the two futures series trading alongside for most of the year. For our NYMEX gasoline data series, we consider unleaded gasoline up to September 1^st^, 2006 and RBOB gasoline thereafter, date at which liquidity for the latter overtook that for the former.

@@

On the one hand, futures markets have a long history of serving commodity producers to allievate their commodity price and output risks. @keynes_treatise_1930 and @hicks_value_1939 emphasise the "normal" behaviour of naturally short hedgers outsourcing their commodity risk to naturally long commodity specialist speculators. This market configuration is referred to as "backwardation" where futures is expected to trade at a discount to spot, providing speculators with incentives to take the long side. The opposite market paradigm is referred to as "contango". Working [@working_theory_1948; @working_hedging_1953] and Cootner [@cootner_returns_1960; @cootner_speculation_1967] somehow relax this "supply-of-speculative-services" [@till_long_2007] approach by introducing processors and merchants, also commodity experts, on the hedgers side and relating the notions of backwardation and contango more directly to the term structure via the theory of storage where the shape of the front end of the term structure (basis) is directly related to the market price of storage.  
Financialization, on the other hand, refers to the entry of financial investors into the commodity futures markets who, for the most part, are not commodity experts. Besides, these new market participants tend to exhibit herding like behaviour, often taking massive long-only positions on the whole cross-section of indexed commodities in an attempt to hedge against inflation and/or seek further diversification benefits for portfolios with, in most cases, existing large positions across various asset classes [@brunetti_speculators_2016; @boyd_prevalence_2016; @cheng_convective_2014; @juvenal_speculation_2015; @singleton_investor_2013; @tang_index_2012]. This contrasts with the traditional approaches of Keynes and Working and in this context the issues of co-movement and aggregate market participants behaviour seem particularly relevant.  
Asset pricing techniques are well suited to study co-movement. Building on theory we develop an ad-hoc commodity based asset pricing framework that we believe is well adapted to study these two issues simultaneously and implement it over four periods of interest. The first, the pre-financialization period, starts in 1997 and is naturally followed by financialization, spanning the 2004/2007 period, and the financial crisis and its aftermath, spanning the 2008/2013 period. The last period spans the 2014/present period sarting with the beginning of the tapering/end of the US Quantitative Easing program (QE).  
Using the weekly price, open interest and market position data described above we construct market portfolios of commodity nearby futures as well as mimicking portfolios for market, commercial hedging pressure, term structure and liquidity risk factors in returns. While our market portfolios are long only, equally weighted portfolios of particular commodity sets, our other mimicking portfolios for risk factors are constructed as combinations of two sub-portfolios (legs), one held long, one held short, which respective constituents are pooled on a weekly basis according to risk factor specific criterias. The weekly return for a particular mimicking portfolio is calculated as the difference between the average return for constituents of the long portfolio and that for the constituents of the short portfolio. i.e, return on the mimicking portfolio for risk factor $i$ ($f_{i}$) for week $t$:

@@

The same operation is repeated for the two ($p=2$) and three ($p=3$) factor model cases, where regressors include in turn all combinations of any two and three risk factors respectivly, as well as for the four ($p=4$) factor model case where all the risk factor mimicking portfolios are included. Results are reported for the one-factor model case with the other results available from the authors upon request.  
For each period we also implement the analysis described above over periods of low and high aggregate CHP (**CHP**), a notion which we introduce in this study and which we believe is particularly relevant in the context of financialization where issues of co-movement and aggregate market participants' behaviour play a central role. We define **CHP** as the cross-sectional average of CHPs for a particular pool of assets. i.e, Week $t$ **CHP** for asset pool $x$:

@@

We further define aggregate backwardation (**backwardation**) and contango (**contango**) **CHP** regimes as, for a given period, **CHP** levels below and above the period's median respectively and, in contrast with [@hong_what_2012] who aggregate across sub-sectors, the asset pool we consider for **CHP** construction include the whole cross-section of US traded commodities.

@@

For each period and **CHP** regime we record the proportion of time each pick is held in the long and (/or) short leg of the corresponding risk factor mimicking portfolio as well as the average pairwise correlation amongst every sets of picks.

@@

# Results & discussion {#results}

The descriptive statistics over the various time periods for our two sets of commodities (US and non-US) as well as the four factors considered are shown in Table \ref{table1}. The traditional Keynesian hedging pressure paradigm postulates a negative relationship between the return on an individual commodity and its hedging pressure. Financialization has been seen as a cross-sectional commodity market phenomenon; this naturally leads us to extend the individual hedging pressure paradigm by incorporating aggregate CHP (**CHP**), a market wide measure of hedging pressure. The extended Keynesian hedging pressure paradigm should also imply that returns for individual commodities should be high in periods where **CHP** is low and vice et versa.  
In table \ref{table2} the mean returns are shown over high (**contango**) and low (**backwardation**) **CHP** regimes for each time period. Over the first period eighteen of the twenty-four US commodities have higher mean returns during periods of **backwardation** as compared to **contango**, with the difference being statistically significant at the 5% level for one commodity and for the remaining six commodities the higher mean return during **contango** is not statistically significant for any of them. For the six UK commodities, all metals, four of them had higher mean returns during phases of **backwardation** as compared to **contango** in the first period, with one of these differences being statistically significant at the 5% level while for the remaining two the higher mean return during **contango** is not statistically significant for either. An equally weighted portfolio of the twenty-four US commodities had a mean return of 10.98% during **backwardation** phases, statistically significant at the 1% level, and a mean return of 0.93% during **contango**, both in the first period, with the difference between them statistically significant at the 5% level. The corresponding figures for an equally weighted portfolio of the UK metals were 8.35% and -1.71%, the difference again statistically significant at the 5% level. These results provide broad support for an aggregate version of the Keynesian hedging pressure hypothesis and suggests that hedgers in the aggregate were engaged in risk transfer during this period.  
Over the second period the pattern essentially reverses. For the twenty-four US commodities over this period, eighteen have higher mean returns during phases of **backwardation** over **contango**, with the difference statistically significant for four commodities and of the remaining seven the higher mean return during **backwardation** is statistically significant for one commodity. All of the six UK metals achieved higher mean returns over **backwardation** than contango with the difference being statistically significant for one commodity. The equally weighted US portfolio had a mean return of 11.16% over **backwardation** while it was 26.68% during **contango**, the first statistically significant at the 5% level and the second at the 1% level, with the difference statistically significant at the 1% level. The difference was considerably more pronounced for the equally weighted UK metals portfolio with a mean return of 6.43% during **backwardation** rising to 41.22% during **contango}, with the difference also significant at the 1% level.  
Thus the onset of financialization seems to have engendered a change in the nature of hedger’s behaviour in the aggregate, taking it away from risk shifting, the traditional Keynesian view. This phenomenon has been noted in earlier theoretical and empirical work [@danthine_information_1978, @stout_why_1998] in different context from which two competing models of behaviour have emerged. The information arbitrage model implies that hedgers may often be seeking out counterparties to trade with rather than being purely passive. This issue is also raised in both @cheng_financialization_2014 and @stulz_rethink_1996 who point out that hedgers may be taking a view on prices just as speculators do. In fact, by hedging away some of their risk, hedgers are able to speculate more heavily based on their disagreement against speculators regarding futures price movement [@simsek_speculation_2013], which fits into a heterogenous expectation theory of speculation. As @stout_why_1998 points out, this disagreement-based theory of trading is one of the main reason the public and the law disapproves the speculators as this form of trading is regarded as non-productive^[@duffie_challenges_2014 also discusses some of the challenges faced by a policy treatment of speculative trading motivated by differences in beliefs.]. In the second model hedgers are seen as liquidity providers as in @kang_tale_2017. This explanation is also consistent with the nature of financialization which led to the arrival of long term, long only investors who require substantial liquidity to roll-over their positions.

The average time series pricing performance of the various one factor models on the US traded assets over the four periods is shown in Table \ref{table3}. Over the first period the market factor has the best average time series performance with the highest average adjusted R^2^ with all three futures market factors performing poorly and the overall pricing performance consistent across periods of high and low **CHP** (below period median). The two factor models average R^2^s (results available upon request) are roughly the sum of the one factor R^2^s indicating that the factors are almost orthogonal over this period. The onset of financialization leads to a substantial improvement in the performance of the CHP factor with its average R^2^ increasing to 5.4% in the second period relative to 1.4% in the first. The performance of the market factor also improves though not by as much in a relative sense (18.6% from 12.2%). Both the term structure factor and the open interest growth factor show a much smaller improvement in pricing performance. Over all periods, all of our factor models show similar pricing performance across regimes of high and low **CHP}.

@@

kable(table, align = c(rep('l', 2L), rep('r', 4L)), digits = getOption("digits"),
      booktabs = TRUE, longtable = TRUE, linesep = "", caption = caption) %>%
  kableExtra::landscape() %>% 
  kableExtra::kable_styling(full_width = TRUE, font_size = 10L) %>%
  kableExtra::column_spec(1L, width = "5cm") %>%
  kableExtra::column_spec(2L, width = "2cm")

@@

kable(table, align = c(rep('l', 3L), rep('r', 4L)), digits = getOption("digits"), 
      booktabs = TRUE, longtable = TRUE, linesep = "", caption = caption) %>% 
  kableExtra::landscape() %>% 
  kableExtra::kable_styling(font_size = 10L, full_width = TRUE) %>% 
  kableExtra::column_spec(1L, width = "3cm") %>% 
  kableExtra::column_spec(2L, width = "3cm") %>% 
  kableExtra::column_spec(3L, width = "2cm")
```

@@

kable(table, align = c(rep('l', 2L), rep('r', 4L)), digits = getOption("digits"), 
      booktabs = TRUE, longtable = TRUE, linesep = "", caption = caption) %>% 
  kableExtra::landscape() %>% 
  kableExtra::kable_styling(font_size = 10L, full_width = TRUE) %>% 
  kableExtra::column_spec(1L, width = "5cm") %>% 
  kableExtra::column_spec(2L, width = "2cm")
```
